---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tintin <img src="man/figures/logo.svg" align="right" height="139" alt="" />

<!-- badges: start -->
<!-- badges: end -->

The goal of tintin is to provide palettes generated from Tintin covers. There is one palette per cover, with a total of 24 palettes of 5 colours each. Includes functions to interpolate colors in order to create more colors based on the provided palettes.

```{r pallettes, echo=FALSE, fig.width=7, fig.height=7}
showcol <- function (pal = rev(tintin::tintin_colours), n = NULL, type = "all", exact.n = TRUE) {
  gaplist <- ""

  maxnum <- 5

  palattr <- switch(type, qual = "qualitative", div = "divergent",
                    seq = "sequential", all = "qualitative+divergent+sequential")
  if (is.null(n))
    n <- maxnum

  if (length(n) == 1)
    n <- rep(n, length(pal))

  n[n < 3] <- 3
  n[n > maxnum] <- maxnum[n > maxnum]
  nr <- length(pal)
  nc <- max(n)
  ylim <- c(0, nr)
  oldpar <- par(mgp = c(2, 0.25, 0))
  on.exit(par(oldpar))
  plot(1, 1, xlim = c(-4, nc), ylim = ylim, type = "n", axes = FALSE,
       bty = "n", xlab = "", ylab = "")
  for (i in 1:nr) {
    nj <- n[i]
    if (pal[i] == "")
      next
    shadi <- pal[[i]]
    rect(xleft = 0:(nj - 1), ybottom = i - 1, xright = 1:nj,
         ytop = i - 0.2, col = shadi, border = "light grey")
  }
  text(rep(-0.1, nr), (1:nr) - 0.6, labels = names(pal), xpd = TRUE,
       adj = 1)
}

showcol()
```

## Installation

You can install the development version of tintin like so:

``` r
remotes::install_github("pachadotdev/tintin")
```

## Example

This is a basic example which shows you how to create a plot. We'll plot
the top five causes of injury in the `tintin_head_trauma` dataset that comes 
with the package.

```{r example1}
library(dplyr)
library(ggplot2)
library(tintin)

total_head_trauma_5 <- tintin_head_trauma %>% 
  arrange(-loss_of_consciousness_length) %>% 
  filter(row_number() <= 5)

ggplot(total_head_trauma_5) +
  geom_col(aes(x = cause_of_injury, y = loss_of_consciousness_length, 
    fill = book_title), position = "dodge") +
  labs(x = "Cause of injury", y = "Loss of consciousness length",
    title = "Top five causes of injury") +
  theme_minimal() +
  scale_fill_manual(values = tintin_colours$the_black_island,
    name = "Book") +
  coord_flip()
```

What is special about the package is being able to pass the colours as a 
function to `ggplot2`. We'll adapt the previous example to show that case.

```{r example2}
ggplot(total_head_trauma_5) +
  geom_col(aes(x = cause_of_injury, y = loss_of_consciousness_length, 
    fill = book_title), position = "dodge") +
  labs(x = "Cause of injury", y = "Loss of consciousness length",
    title = "Top five causes of injury") +
  theme_minimal() +
  scale_fill_tintin_d(option = "cigars_of_the_pharaoh", direction = -1) +
  coord_flip()
```

```{r example2_2, eval = FALSE}
# Note that I can also write the palette name as
# "cigars of the pharaoh" or even as "CiGaRS  of ThE Pharaoh"
ggplot(total_head_trauma_5) +
  geom_col(aes(x = cause_of_injury, y = loss_of_consciousness_length, 
    fill = book_title), position = "dodge") +
  labs(x = "Cause of injury", y = "Loss of consciousness length",
    title = "Top five causes of injury, again") +
  theme_minimal() +
  scale_fill_tintin_d(option = "cigars of the pharaoh", direction = -1) +
  coord_flip()
```

What happens if we need more colours than 5? The functions in the package
can fix that. We'll plot the top ten causes of injury.

```{r example3}
total_head_trauma_10 <- tintin_head_trauma %>% 
  arrange(-loss_of_consciousness_length) %>% 
  filter(row_number() <= 10)

ggplot(total_head_trauma_10) +
  geom_col(aes(x = cause_of_injury, y = loss_of_consciousness_length, 
    fill = book_title), position = "dodge") +
  labs(x = "Cause of injury", y = "Loss of consciousness length",
    title = "Top ten causes of injury") +
  scale_fill_manual(values = tintin_clrs(
    n = length(unique(total_head_trauma_10$book_title)), 
    option = "the black island"),
    name = "Book") +
  coord_flip()
```

```{r example3_2, eval = FALSE}
# or alternatively

ggplot(total_head_trauma_10) +
  geom_col(aes(x = cause_of_injury, y = loss_of_consciousness_length, 
    fill = book_title), position = "dodge") +
  labs(x = "Cause of injury", y = "Loss of consciousness length",
    title = "Top ten causes of injury") +
  scale_fill_manual(values = tintin_pal(option = "the black island")(8), 
    name = "Book") +
  coord_flip()
```

The use of colour instead of fill is analogous. Let's plot the top ten causes of
injury per year to see it.

```{r example4}
library(tidyr)

total_head_trauma_y <- tintin_head_trauma %>% 
  group_by(year) %>% 
  summarise_if(is.integer, sum) %>% 
  pivot_longer(loss_of_consciousness_length:loss_of_consciousness_severity) %>%
  mutate(
    name = ifelse(name == "loss_of_consciousness_length", "Length", "Severity")
  )

ggplot(total_head_trauma_y) +
  geom_line(aes(x = year, y = value, color = name), linewidth = 1.5) +
  labs(x = "Year", y = "Intensity",
    title = "Loss of consciousness length and severity") +
  theme_minimal() +
  scale_colour_manual(
    values = tintin_pal(option = "tintin in the land of the soviets")(2), 
    name = "Cause of injury"
  )
```
